---
title: "Penguin Flipper Lengths ANOVA"
author: "Juan Silva"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

packages: tidyverse, car, onewaytests

```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")



```

Our question: is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo and Adelie penguins near Palmer Archipielago?

Single factor is species. Three levels: chinstrap, gentoo and adelie penguins. My continuous variable is flipper length.

Maybe I can use a one-way ANOVA...
- Independent Observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}

counts <- with(penguins, table(species, sex))

# View(counts)

penguins2 <- penguins %>% 
  mutate(
    common_name =
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

View(penguins2)

```


Exloratory graphs and leven's test for equal variances

```{r}


hists <- ggplot(penguins2, aes(x = flipper_length)) + 
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")

hists

qqplot <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
    facet_wrap(~ common_name, scale = "free")

qqplot

# Based on qqplots and histograms, observations for each group appear normally distributed. Also, n>30 for all groups, so means will be normally distributed by CLT.




```

Evaluate variances and perform levene's test for equal variances

```{r}
# If the largest sampe variances is > 4x greater that the smallest sample variance, then these are "close engouh" to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances 
# variances are in mm squared
 # the largest variance (33.1 is defintely not 4x greater thatn the smallest variance (15.2))

#  Levene's test

# leveneTest function in the "car" package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin

#  H0: There's no difference in variances (variances are equal)
#  HA: Variances are NOT equal

#  There is no significant difference in variances across species. (we retain null hypothesis)



```

One-way ANOVA for penguin flipper length by species(common_name)

```{r}

# ANOVA
# H0: Means across all groups are equal
# HA: At least two groups means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)

# Very unlikely that u could have taken these 3 samples taken from populations that have the same mean and taken these differences by random chance. So probably i took these means from populations with different means.



#  Post Hoc testing by Tukey's HSD

# H0: No significatn differnece in pairwise means
# ha: there IS  a signifcant difference in pairwise means 

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

# Reject null hyp in every pairwise comparison


# aov.test function in the onewaytests packe DOES allow in-line references

```

Statement:

Mean flipper length of female pegnuins differ significantly for all species studied (one-way anova F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD $\alpha\ = 0.05.
