---
title: "Penguin Flipper Lengths ANOVA"
author: "Juan Silva"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

packages: tidyverse, car, onewaytests

```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")



```

Our question: is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo and Adelie penguins near Palmer Archipielago?

Single factor is species. Three levels: chinstrap, gentoo and adelie penguins. My continuous variable is flipper length.

Maybe I can use a one-way ANOVA...
- Independent Observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}

counts <- with(penguins, table(species, sex))

# View(counts)

penguins2 <- penguins %>% 
  mutate(
    common_name =
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

View(penguins2)

```

